The Sub-Millimetre User Reduction Facility: an iterative map-maker for \scuba

### 2012-08-23 smurf_paper.tex
missed some of Douglas' comments

The James Clerk Maxwell Telescope is operated by the Joint Astronomy
Centre on behalf of the Science and Technology Facilities Council of
the United Kingdom, the Netherlands Organisation for Scientific
Research, and the National Research Council of Canada. Additional
funds for the contruction of SCUBA-2 were provided by the Canada
Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
funding.

\textit{Is there some kind of grant number that has to be referenced
  for Ed's CANFAR funding?}

@@

\item Locating each candidate step.

\item Determining the widths of the candidate steps.

\item Determine the height of the step. Steps can vary enormously in height.
For instance the step in the left plot of Fig.~\ref{fig:steps1} is
roughly 3 pW in height, whereas the step at sample 5100 in the right hand
plot is only 0.01 pW in height. The height of a step is taken as the
difference in the local median value just before and just after the step.

\item Correcting all subsequent data values by the step height.

\item Once all steps have been corrected, a global offset is added to the
corrected bolometer time stream to ensure that the mean value before and
after step correction is unchanged.

@@

\subsection{Detecting candidate steps}
The first stage in the algorithm is to search the bolometer time stream for
contiguous blocks of samples that have unusually large mean gradients
compared to the neighbouring samples. Each such block of samples is
considered to be a candidate step. The starting and ending time of each
such block is noted. Once all candidate steps have been found for a
bolometer, the algorithm proceeds to the next stage. Candidate steps are
found as follows:

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{steps2.pdf}
\caption{A section of the right hand bolometer from  Fig.~\ref{fig:steps1}
with the median smoothed data overlayed in red.}
\label{fig:steps2}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{steps3.pdf}
\caption{Differences between adjacent values of the median smoothed data
in  Fig.~\ref{fig:steps2}.}
\label{fig:steps3}
\end{figure}

@@

\item The differences for the remaining samples are smoothed
using a mean filter of width 50 samples. The smoothing process fills in small
holes in the array (up to 40 samples), but will not fill in larger holes.
Such larger holes are filled in later by linear interpolation. The
resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
provide an estimate of the local gradient of the underlying noise-free
median-filtered data stream.

@@

\item Blocks of contiguous samples for which the residual difference exceeds
25 times the local RMS of the residual differences are found. Any two
such blocks that are separated by fewer than 100 samples are merged into
a single block. Blocks are ignored if they fail any of the following
tests:

@@

\item The width of the block must be less than 200 samples.

\item The total of all the \snr\ values (the absolute residual difference
divided by the local RMS of the residual differences) in the block must be
at least 25.

\item The total of all \snr\ values in the block must exceed half the value of
the largest single \snr\ value in the block.

@@

The data in Fig.~\ref{fig:steps5} results in two blocks being defined -
one between samples 5037 and 5102 and another between samples 5286 and
5296. Each of these blocks define the start and end of a candidate step.
The first block includes two visible steps, but they have been merged
into a single candidate step due to their small separation (less than 100
samples).

>>>>

The James Clerk Maxwell Telescope is operated by the Joint Astronomy
Centre on behalf of the Science and Technology Facilities Council of
the United Kingdom, the Netherlands Organisation for Scientific
Research, and the National Research Council of Canada. Additional
funds for the contruction of SCUBA-2 were provided by the Canada
Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
funding. Also mention NSERC.

@@

\item locating each candidate step;

\item determining the widths of the candidate steps;

\item determining the height of the step (steps can vary enormously in
height -- for instance the step in the left-hand panel of
Fig.~\ref{fig:steps1} is roughly 3\,pW in height, whereas the step at
sample 5100 in the right-hand panel is only 0.01\,pW in height -- the
height of a step is taken as the difference in the local median value
just before and just after the step);

\item correcting all subsequent data values by the step height;

\item once all steps have been corrected, a global offset is added to the
corrected bolometer time stream to ensure that the mean value before and
after step correction is unchanged.

@@

\subsection{Detecting candidate steps}
The first stage in the algorithm is to search the bolometer time stream for
contiguous blocks of samples that have unusually large mean gradients
compared to the neighbouring samples. Each such block of samples is
considered to be a candidate step. The starting and ending time of each
such block is noted. Once all candidate steps have been found for a
bolometer, the algorithm proceeds to the next stage. Candidate steps are
found as follows.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{steps2.pdf}
\caption{A section of the bolometer signal from the right-hand panel
of Fig.~\ref{fig:steps1} with the median smoothed data overlaid in
red.}
\label{fig:steps2}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{steps3.pdf}
\caption{Differences between adjacent values of the median smoothed
data shown in Fig.~\ref{fig:steps2}.}
\label{fig:steps3}
\end{figure}

@@

\item The differences for the remaining samples are smoothed
using a mean filter of width 50 samples. The smoothing process fills in small
holes in the array (up to 40 samples), but will not fill in larger holes.
Such larger gaps are filled in later by linear interpolation. The
resulting smoothed differences (shown in blue in Fig.~\ref{fig:steps4})
provide an estimate of the local gradient of the underlying noise-free
median-filtered data stream.

@@

\item Blocks of contiguous samples for which the residual difference
exceeds 25 times the local RMS of the residual differences are
found. Any two such blocks that are separated by fewer than 100
samples are merged into a single block. Blocks are only kept if they
pass all of the following tests:

@@

\item the width of the block must be less than 200 samples;

\item the total of all the \snr\ values (the absolute residual difference
divided by the local RMS of the residual differences) in the block must be
at least 25;

\item the total of all \snr\ values in the block must exceed half the value of
the largest single \snr\ value in the block.

@@

The data in Fig.~\ref{fig:steps5} results in two blocks being defined,
one between samples 5037 and 5102, and another between samples 5286
and 5296. Each of these blocks define the start and end of a candidate
step.  The first block includes two visible steps, but they have been
merged into a single candidate step due to their small separation
(less than 100 samples).
